---
title: "Practical Machine Learning Project"
author: "Robert Kerr"
date: "19/09/2014"
output: html_document
---

First we read in the training data, I had edited this in a spreadsheet package to remove columns that were mostly empty or had mostly NA entries.

```{r}
trainingData <- read.csv("/home/r/Downloads/coursera/pml-training-reduced.csv")
```

Now we load the caret package that will be used to preprocess, fit and evaluate the data.

```{r}
library(caret)
```

Next is the partitioning of the data into a training and a test set. I chose to take 75% of the data for the training set.

```{r}
inTrain = createDataPartition(trainingData$classe, p = 3/4)[[1]]
train = trainingData[inTrain,]
test = trainingData[-inTrain,]
```

Then some preprocessing was carried out, I used PCA to reduce the number of variables. I chose to use variables that would account for 95% of the variance. This gave me 24 variables to use a predictors.

```{r}
preproc <- preProcess(train[7:58],method="pca",thresh=0.95)
trainPCA <- predict(preproc,train[7:58])
```

Note that I didn't use columns that could not be used as predictors.

Now we set up a training control object that we can pass into the main training method. This will perform 10 separate 10 fold cross validations

```{r}
trControl <- trainControl(method="repeatedcv", number=10, repeats=10)
```

Now we actually train our model. I've set ntree to be 1 here but when I ran it for the purposes of submission I set ntree to be 50.

```{r}
modFit <- train(train$classe~.,data=trainPCA,method="rf",trControl=trControl,ntree=1)
modFit
```
From this we can expect an accuracy of around 80% though cross validation tends to over estimate this. This would suggest an out of sample error rate of about 20%.

We now run the PCA preprocessing on the testing subset...
```{r}
testPCA <- predict(preproc,test[7:58])
```

...and generate a confusion matrix on this test data and we can see that it is about 82% accurate here, suggesting an out of sample error rate of around 16%, which is close to the rate estimated during the cross validation.

```{r}
confusionMatrix(test$classe,predict(modFit,testPCA))
```

I then loaded the test data provided on the class site and ran the predict method on it, but I won't show that here because it will give my answers away!
